---
title: "A Very Serious Analysis of the Stationarity of Corn and Soybean Prices"
author: "Peter Economist, Paul Economist, Mary Economist"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  pdf_document:
    template: simple.latex
    fig_caption: yes
documentclass: ajae
bibliography: bibliography.bib
---

 



```{r, warning = FALSE, message = FALSE, echo=FALSE}
library(xtable)
library(ggplot2)
library(ggfortify)
library(gridExtra)
source('data-raw/fetch-raw-data.R')
load('analysis-output/results.Rda')
```


# Introduction 

This paper is a very serious examination of the relationship between the December 2016 corn and November 2016 soybean futures prices.

# Data

We use daily settlement prices for the December 2016 and November 2016 corn and soybeans futures contracts on the Chicago Board of Trade from January 2, 2015 to `r format(Sys.time(),"%B %d, %Y")`. The data are plotted in Figure 1. 

# Methods 

We conduct a thorough statistical analysis, begining with pretesting for stationarity and cointegration; then we proceed to fitting an econometric model of the price relationships. 

## Pre-testing for Stationarity and Cointegration

We conduct an adf test on the corn and sobean price series to determine if they have a unit root [@said1984testing, @enders1995applied].

\begin{equation}
\Delta y_t = \alpha + \gamma y_{t-1} + \delta_1 \Delta y_{t-1} + \dots + \delta_{p-1} \Delta y_{t-p-1} + \epsilon_t
\end{equation}

where $p$ is the order of the autogressive term in the prices. The test will determine if there is a unit root, H0: $\gamma = 0$, with no drift, H0: $\alpha = 0$. 

# Results

We found there to be a unit root in soybeans, but corn exhibited stationarity. The values of the ADF statistics for corn and soybeans were `r round(adf[[1]]@teststat[1], digits = 2)` and `r round(adf[[2]]@teststat[1], digits = 2)`, respectively.



\clearpage


```{r, echo=FALSE, warning = FALSE, message = FALSE, results = "asis"}
t = list()
t[[1]] <- xtable(adf[[1]]@testreg, caption = "ADF Results for Corn")
t[[2]] <- xtable(adf[[2]]@testreg, caption = "ADF Results for Soybeans")
print.xtable(t[[1]], caption.placement = 'top', comment = FALSE)
```

```{r, echo=FALSE, warning = FALSE, message = FALSE, results = "asis"}
print.xtable(t[[2]], caption.placement = 'top', comment = FALSE)
```

```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.cap = 'Corn and Soybean Prices' }
plot1 <- autoplot(CZ2016, ylab = "Corn Prices, cents/bu") + theme_bw()
plot2 <- autoplot(SX2016, ylab = "Soybean Prices, cents/bu") + theme_bw()
grid.arrange(plot1, plot2, ncol=1)
```



\clearpage

# References 
