---
title: "Reproducible Research Project Template"
author: 'Mindy L. Mallory'
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  md_document:
    toc: true
bibliography: bibliography.bib
---


# Research Project Template

This repository contains a template for a reproducible research project. The fundamental idea of reproducible research is that the steps that take your research from raw data to manuscript, thesis, or report should be fully automated. This way, your work can be checked by your adviser, mentors, collaborators, others working in your area, journal reviewers, and your future self. 

I became interested in reproducible research because I was tired of being terrified of my own analysis. I was constantly petrified someone would question my work and ask me to open the black box and verify what I did was correct; in some cases even being asked reproduce a result was terrifying because I knew the convoluted path of data prep and cleaning that I took in arriving at my result. 

I read the book [Reproducable Research with R and RStudio](http://www.amazon.com/Reproducible-Research-Studio-Chapman-Hall/dp/1466572841) by Christopher Gandrud [-@gandrud2013reproducible], and I read a lot of blog posts and tutorials by [Karl Broman](http://kbroman.org/pages/software.html) and [Carl Boettiger](http://www.carlboettiger.info/2012/05/06/research-workflow.html) and I struck out on my own path to execute a reproducible research project from start to finish. The repository for that project is [here](https://github.com/mindymallory/BBOBAS), while I was successful in learning the basics of how r, rmarkdown, knitr, and pandoc combine to make reproducible research possible, you can tell just by looking at the project's Github repository that I failed miserably at making the project reproducible. This is because the repository is totally unorganized, and I am pretty sure I am the only one who could reproduce the results from this project. 

But in that failure, I learned a lot about how a reproducible research project should be organized. I built this template for my future students and for my future self. 

R and RStudio are an excellent vehicle for conducting reproducible research. You write manuscripts and reports in `.rmarkdown` documents that includes code chunks that perform analysis. The code chunks are evaluated by `R` and incorporated in the document by the tools in the `knitr` package to produce a markdown `.md` document. From there a program called `pandoc` converts your markdown document to whatever file format you like: PDF (formatted with latex .csl files), html, or Microsoft Word. This all happens without the user really knowing what is going on, which makes it easy to get started.    

# Getting Started

First download the repository to your local machine. If you use Github, then this will mean cloning the repository into a new R project. If you are not a Github user, simply click 'Download Zip' and extract the file to a convenient location. 

## Repository Contents

The repository contains several folders and files. They are organized to keep data preparation and cleaning in one file, `data-raw`, analysis in another, `analysis`, and outputs from the analysis that will become tables and figures, and numbers in the text of the manuscript, `analysis-output`.  The remaining files in the root directory are files related to the manuscript itself. Next, we'll demonstrate how to link the raw data to the analysis to the output to the manuscript so that all the steps to generate the manuscript are automated and thus, not subject to the inconsistencies that go along with piece-wise data preparation and analysis.

# Generate the Manuscript with One Click

Open the `manuscript-example.Rmd` and `tablesandfigures-examples.Rmd` files in RStudio. Click the 'Knit PDF' button on the code editing pane. Install the following packages, if they are not already installed:

```
install.packages(xtable)
install.packages(ggplot2)
install.packages(ggfortify)
install.packages(gridExtra)
install.packages(Quandl)
install.packages(RCurl)
install.packages(xts)
install.packages(urca)
install.packages(vars)
```

Click the 'Knit PDF' button on the code editing pane, and voila! A PDF of the manuscript should appear. In what follows we will walk through what is happening step by step. 

# Putting it all Together

Now we see the contents of the data and analysis files and how they come together. 

## The `data-raw` Folder

The `data-raw` folder should either contain your raw data files (that will **never** *ever* be modified), or a script that makes and api call, or pulls the raw data in from a shared server, etc. In this example file, there is a script called `fetch-raw-data.R`, and its contents are shown below. This file fetches corn and soybean price data from [quandl.com](quandl.com) and puts them in data objects called `CZ2016 and SX2016`. Then it converts the data to `xts` objects, and trims the dates to the study period of interest. 

```{r, eval= FALSE, warning = FALSE, message = FALSE}
# Filename: fetch-raw-data.R
# This file fetches the raw data and performs pre-processing (cleaning) to get it ready for analyzs

library(RCurl)
library(xts)
library(Quandl)
Quandl.api_key("79SfoMaQc1npRAuq9ExZ")
# Define Dates of Analysis
  start  <- '2015-01-01'
  today  <- format(Sys.time(),"%Y-%m-%d")

# Fetch Corn and Soybean Prices
  CZ2016 <- Quandl("CME/CZ2016", type = "xts")
  SX2016 <- Quandl("CME/SX2016", type = "xts")

# Trim the dates
  CZ2016 <- CZ2016[paste0(start,'/',today), 'Settle']
  SX2016 <- SX2016[paste0(start,'/',today), 'Settle']

```

Of course, every data cleaning and preparation activity will be different, but in this file you should do all the preparation so that the objects created by this script are ready to be accepted in the `analysis.R` script. 

## The `analysis` Folder

The contents of the `analysis` folder are below. The key is the line that says, `source('data-raw/fetch-raw-data.R')`. This calls the `fetch-raw-data.R` script so that when you run the code below, the raw data are fetched and prepared (from scratch each time you run the script). Then, the following contents of the `analysis.R` script test the corn and soybean prices for the presence of unit roots via the ADF test [@said1984testing]. 




```{r, warning = FALSE, message = FALSE}


# Filename: analysis.R
# This file performs statistical analysis. It could be just one file, so it doesn't neccessarly 
#  need it's own folder, but sometimes your analysis may get complicated enough that you want 
# to compartmentalize it. Separating different types of analyses into different scripts contained
#  in the same folder can facilitate this

library(urca)
library(vars)
# This line runs the source code that fetched your raw data and cleaned it. Now it is available 
# for conducting analysis.
source('data-raw/fetch-raw-data.R')

# Store results of ADF tests for Corn and Soybeans in a list
adf      <- list()
adf[[1]] <- ur.df(CZ2016, type = 'drift', lags = 5)
adf[[2]] <- ur.df(SX2016, type = 'drift', lags = 5) 


# Store results of a Johansen cointegration test for Corn and Soybeans 
jct      <- ca.jo(cbind(CZ2016, SX2016), type = 'eigen', K = 5)


# Fit a VAR

lag_selection <- VARselect(cbind(CZ2016, SX2016), lag.max = 8)

var_model <- VAR(cbind(CZ2016, SX2016), p = 1, type = "const")

# Save these results so that it can be pulled into the manuscript without re-running analysis.
save(adf, jct, lag_selection, var_model, file = 'analysis-output/results.rda')



```

## The `analysis-output` Folder

The last line of the code snippet above says, `save(adf, jct, lag_selection, var_model, file = 'analysis-output/results.rda')`. What this does is save the objects that contain the adf, Johansen cointegration, and VAR regression results to an `.rda` file called `results.rda` in the `analysis` folder. This 'R Data' file can be read in by R and the variable names, `adf`, `jct`, `lag_selection`, and `var_model` are preserved when loaded later. We will load the `results.rda` file into the `tablesandfigures-example.Rmd` document to make tables and figures in the manuscript. 

## The Manuscript

At the top of the file named `manuscript-example.Rmd` you see a YAML (Yet Another Mark Up Language) header. This header tells knitr and pandoc what exactly you want done with the document. 

    ---
    title: "A Very Serious Analysis of the Stationarity of Corn and Soybean Prices"
    author: "Peter Economist, Paul Economist, Mary Economist"
    date: '`r format(Sys.time(), "%B %d, %Y")`'
    output: 
      pdf_document:
        template: simple.latex
        fig_caption: yes
    documentclass: ajae
    bibliography: bibliography.bib
    ---

`Title` and `author` are self explanatory.  

`date`: field tells knitr to place the current date formatted in the `%B %d, %Y` style.  

`output`: After knitr evaluates code chunks contained in the body of the file. The output feild tells pandoc what kind of file to create. Here we have specified to produce PDF output. PDF output is produced by pandoc creating a `.tex` file and if no further fields are specified there is a latex template that pandoc uses to make the docuement (based on the `article` class). Here we have specified to create the manuscript according to the specifications of the American Journal of Agricultural Economics (AJAE). Since they have their own latex class (`ajae`) that comes in the standard latex distribution we can just specify `documentclass: ajae` and the formatting is handled. We needed to also specify `template: simple.latex` because something in the pandoc template was clashing with the `ajae.csl` file. I removed the problem lines and saved that as `simple.latex`, which you can see in the root directory of this repository. We will cover how to specify different output formats in a later section. 

`bibliography`: The file `bibliography.bib` is located in the root directory of this repository and it is a Bibtex database of all the references needed for the manuscript. Open this file and note what the reference entries look like. To build a database for your own paper, Google Scholar has a 'cite' button below every search result it returns. Click 'cite', then click 'Bibtex' and a plain text window will open with the properly formatted Bibtex entry. Just copy and paste this into `bibliography.bib`. 

### Markdown Basics

Formatting a docuemnt with Markdown is very easy and there are many resources to learn the basics. Start with [http://rmarkdown.rstudio.com/index.html](http://rmarkdown.rstudio.com/index.html) and explore. 

Main points: 

    # This is a Level 1 Header
    
# This is a Level 1 Header

    ## This is a Level 2 Header
    
## This is a Level 2 Header

    This is a citation of Akerlof's Lemons paper [@akerlof1970vthe].
    
This is a citation of Akerlof's Lemons paper [@akerlof1970vthe].

### Code Chunks

This is an example of a code chunk that is in the manuscript document. The opening line tells `knitr` that what follows is code chunk to be evaluated. 


     ``{r, echo=FALSE, warning = FALSE, message = FALSE, results = "asis"}
     t = list()
     t[[1]] <- xtable(adf[[1]]@testreg, caption = "ADF Results for Corn")
     t[[2]] <- xtable(adf[[2]]@testreg, caption = "ADF Results for Soybeans")
     print.xtable(t[[1]], caption.placement = 'top', comment = FALSE)
     ```



# References
